Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-15 ICML-submission/from-weak-separability-to-strong-separability.tex
.

Consider the
classical real separable Hilbert space $\ell_2 = \{ x \in \R^\infty ~:~
\sum_{i=1}^\infty x_i^2 < + \infty \}$ equipped with the standard inner product
$\ip{x}{x'}_{\ell_2} = \sum_{i=1}^\infty x_i x'_i$.
Through indexing the coordinates of
$\ell_2$ by $d$-tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ of non-negative
integers, the feature map that corresponds to $k$, $\phi: \B(0,1) \to \ell_2$, is defined as follows:
\begin{align*}
& \left(\phi(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \\
& \qquad \cdot \sqrt{2^{-(\alpha_1 + \alpha_2 + \dots + \alpha_d)} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\end{align*}
where $\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d} = \frac{(\alpha_1 + \alpha_2 + \dots + \alpha_d)!}{\alpha_1! \alpha_2! \dots \alpha_d!}$
is the multinomial coefficient.
It can be easily checked that
\[ k(x,x') = \ip{\phi(x)}{\phi(x')}_{\ell_2}. \]

@@

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in \R^d \times \{1,2,\dots,K\}$
labeled examples that are weakly linearly separable with margin $\gamma > 0$
such that $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le 1$.
Let
\begin{align*}
r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1 \quad \text{and} \quad s = \left \lceil \log_2(2/\gamma) \right \rceil \\
\gamma_1 & = \frac{1}{2\sqrt{K}}  \\
& \ \cdot \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{-\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \\
\gamma_2 & = \frac{2^{s(s+1)r(K-1)} }{4\sqrt{K}(2K-5/2) 2^{K-1}} \\
& \ \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)} \; .
\end{align*}
Then, the labeled examples $(\phi(x_1), y_1), (\phi(x_2), y_2), \dots,
(\phi(x_T), y_T)$ are strongly linearly separable with margin $\gamma' =
\max\{\gamma_1, \gamma_2\}$ and $\norm{\phi(x_1)}_{\ell_2},
\norm{\phi(x_2)}_{\ell_2}, \dots, \norm{\phi(x_T)}_{\ell_2} \le 2$.
\end{theorem}

We derive a mistake bound for Algorithm~\ref{algorithm:kernelized} with kernel
function $k(x,x') = \ip{\phi(x)}{\phi(x')}_{\ell_2}$ by simply invoking
\autoref{theorem:strongly-separable-examples-mistake-upper-bound} with the
modified margin $\gamma'$ from \autoref{theorem:margin-transformation} below.
The mistake bound is stated as \autoref{corollary:weakly-separable-examples-mistake-upper-bound}.

As a corollary of \autoref{theorem:margin-transformation} and
\autoref{theorem:strongly-separable-examples-mistake-upper-bound}
we get the mistake bound for Algorithm~\ref{algorithm:kernelized}
under the weak separability condition.

\begin{corollary}[Mistake upper bound]
\label{corollary:weakly-separable-examples-mistake-upper-bound}
Let $K$ be a positive
integer, $\gamma$ be a positive real number, $R$ be a non-negative real number. If
$(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is a sequence of labeled examples in
$V \times \{1,2,\dots,K\}$ such that
\begin{enumerate}
  \item the examples are weakly linearly separable with margin $\gamma$ (Definition~\ref{definition:weak-linear-separability}),
  \item $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le 1$,
\end{enumerate}
then the expected number of mistakes
Algorithm~\ref{algorithm:kernelized}
makes is at most $2^{\widetilde{O}(\min(K \log^2
\frac{1}{\gamma}, \sqrt{\frac{1}{\gamma}} \log K))}$.
\end{corollary}

\subsection{Proof of Theorem~\ref{theorem:margin-transformation}}
The idea behind construction and analysis of the mapping $\phi$ is polynomial
approximation. According to the well known Stone-Weierstrass theorem (see
e.g.~\citet[Section~10.10]{Davidson-Donsig-2010}), on a compact set,
multivariate polynomials uniformly approximate any continuous function.
Intuitively speaking, we use a multivariate polynomial to approximate, on the
unit ball of $\R^d$, the indicator function corresponding to the intersection of
$m=K-1$ halfspaces. Within margin $\gamma$ along the decision boundary, we allow
the polynomial to attain arbitrary value. The polynomial separates examples in
one class from examples in the other classes. To be able to quantify $\gamma'$
we need to quantify certain parameters of the approximating polynomial. We
construct two different polynomials with different parameters. The parameters
are quantified in
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
stated below.

>>>>

Consider the
classical real separable Hilbert space $\ell_2 = \{ x \in \R^\infty ~:~
\sum_{i=1}^\infty x_i^2 < + \infty \}$ equipped with the standard inner product
$\ip{x}{x'}_{\ell_2} = \sum_{i=1}^\infty x_i x'_i$.
Through indexing the coordinates of
$\ell_2$ by $d$-tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ of non-negative
integers, the feature map that corresponds to $k$, $\phi: \B(0,1) \to \ell_2$, is defined as follows:
\begin{align}
& \left(\phi(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \nonumber \\
& \qquad \cdot \sqrt{2^{-(\alpha_1 + \alpha_2 + \dots + \alpha_d)} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\label{eqn:phi}
\end{align}
where $\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d} = \frac{(\alpha_1 + \alpha_2 + \dots + \alpha_d)!}{\alpha_1! \alpha_2! \dots \alpha_d!}$
is the multinomial coefficient.
It can be easily checked that
\[ k(x,x') = \ip{\phi(x)}{\phi(x')}_{\ell_2}. \]

@@

The following theorem is our main technical result in this section; we defer its
proof to Section~\ref{sec:margin-trans}.

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_T, y_T)$ in $\R^d \times \{1,2,\dots,K\}$
be a set of labeled examples that is weakly linearly separable with margin $\gamma > 0$.
In addition, for all $t$ in $\cbr{1,\ldots,T}$, $\norm{x_t} \le 1$.
Let $\phi$ be as defined in Equation~\eqref{eqn:phi}, and
\begin{align*}
r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
\gamma_1 & = \frac{1}{2\sqrt{K}}  \\
& \ \cdot \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{-\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}, \\
\gamma_2 & = \frac{2^{s(s+1)r(K-1)} }{4\sqrt{K}(2K-5/2) 2^{K-1}} \\
& \ \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)} \; .
\end{align*}
Then, the set of labeled examples transformed by $\phi$,
namely $(\phi(x_1), y_1), (\phi(x_2), y_2), \dots,
(\phi(x_T), y_T)$, is strongly linearly separable with margin $\gamma' =
\max\{\gamma_1, \gamma_2\}$. In addition, $\norm{\phi(x_1)}_{\ell_2},
\norm{\phi(x_2)}_{\ell_2}, \dots, \norm{\phi(x_T)}_{\ell_2} \le 2$.
\end{theorem}

We are now ready to derive a mistake bound for Algorithm~\ref{algorithm:kernelized} with kernel
function $k(x,x') = \ip{\phi(x)}{\phi(x')}_{\ell_2}$. This is done by simply invoking
\autoref{theorem:kernelized-upper-bound} with the
modified margin $\gamma'$ from \autoref{theorem:margin-transformation}.
The mistake bound is stated as \autoref{corollary:weakly-separable-examples-mistake-upper-bound}.

%As a corollary of \autoref{theorem:margin-transformation} and
%\autoref{theorem:strongly-separable-examples-mistake-upper-bound}
%we get the mistake bound for Algorithm~\ref{algorithm:kernelized}
%under the weak separability condition.

\begin{corollary}[Mistake upper bound]
\label{corollary:weakly-separable-examples-mistake-upper-bound}
Let $K$ be a positive
integer, $\gamma$ be a positive real number, $R$ be a non-negative real number. If
$L = ((x_1, y_1), (x_2, y_2), \dots, (x_T, y_T))$ is a sequence of labeled examples in
$V \times \{1,2,\dots,K\}$ such that
\begin{enumerate}
  \item $L$ is weakly linearly separable with margin $\gamma$,
  \item for all $t$ in $\cbr{1,2,\ldots,T}$, $\norm{x_t} \le 1$,
\end{enumerate}
then the expected number of mistakes
Algorithm~\ref{algorithm:kernelized}
makes is at most $2^{\widetilde{O}(\min(K \log^2
\frac{1}{\gamma}, \sqrt{\frac{1}{\gamma}} \log K))}$.
\end{corollary}
%, \norm{x_2}, \dots, \norm{x_T}
% (Definition~\ref{definition:weak-linear-separability})

\subsection{Proof of Theorem~\ref{theorem:margin-transformation}}
\label{sec:margin-trans}
The idea behind construction and analysis of the mapping $\phi$ is polynomial
approximation. According to the well known Stone-Weierstrass theorem (see
e.g.~\citet[Section~10.10]{Davidson-Donsig-2010}), on a compact set,
multivariate polynomials uniformly approximate any continuous function.
Intuitively speaking, we use a multivariate polynomial to approximate, on the
unit ball of $\R^d$, the indicator function corresponding to the intersection of
$m=K-1$ halfspaces. Within margin $\gamma$ along the decision boundary, we allow
the polynomial to attain arbitrary value. The polynomial separates examples in
one class from examples in the other classes. To be able to quantify $\gamma'$
we need to quantify certain parameters of the approximating polynomial. We
construct two different polynomials with different parameters. The parameters
are quantified in
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
stated below.
